---
title: "Optimal Duration of Systemic Steroids Use in COVID-19 Treatment: A systematic Review and Meta-analysis"

author: ''
date: "12/09/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE,warning=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE,echo=FALSE}
library(jtools)
library(tmap)
library(meta)
library(metafor)
library(RISmed)
library(reticulate)
library(dplyr)
library(tibble)
library(stringr)
library(tidytext)
library(janeaustenr)
library(magrittr)
library(ggplot2)
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
library(readxl)
library(mgcv) 
library(mgcViz)
library(readxl)
library(mediation)
library(lme4)
library(survminer)
library(survival)

library(sjPlot)
library(sjlabelled)
library(sjmisc)

  
  
```

Question: What is the optimal duration for the use of steroids in COVID-19 treatment?

Objectives: 1) Overall association of systemic steroid use and in-hospital mortality
            2) Effect Modification by Time (Continous, Dichotomy (Binary))
            3) Other SE (Hyperglycemia, Infection)


```{r, echo=FALSE}

rm(list=ls())
#dat=read.csv("steroids.csv")
#data<- read_excel("steroids3.xlsx")
data=read.csv("steroids7.csv")
```









```{r,echo=FALSE,warning=FALSE,fig.height=5, fig.width=8}

library(ggrepel)
library(colorspace)
library(colorblindr)
library(meta)
library(openxlsx)
library(splines)

knots = quantile(data$mean_day, p = c(0.25, 0.5, 0.75), na.rm = T)
f3 <- data %>%

   #filter(Country != "Qatar") %>%
  #arrange(desc(N_vaccinated)) %>%
  ggplot(aes(x = mean_day, y = RR)) +
  geom_smooth(method = "lm", formula = y ~ ns(x,knots = knots),col="red", lty=2, se=T)+
  #ggplot(aes(x = Period, y = Rates, fill=Period))+  
  #geom_smooth(se = T, method = "loess", size = 1.5) +
  #geom_boxplot(show.legend = F)+
  #geom_point(aes(fill = Vaccine, size=N_vaccinated), shape = 21, alpha = .7) +
  expand_limits(x=c(0,20), y=c(0, 4))+
  #colorblindr::scale_fill_OkabeIto(name = "Vaccine type") +
  #scale_fill_jama(name = "Vaccine type")+
  scale_size(name = "",
             range = c(2, 10), breaks = c(10000000, 500000000)) +
 # range = c(2, 10), breaks = c(10000, 500000, 1000000, 1500000, 2000000, 2500000, 4000000)) +
  #annotate("text", label = "p==0.33~~R^{2}==0.00", x = 120, y = 95, parse = T) +
  geom_text_repel(aes(label = Country), size = 2.5) +
  facet_wrap(~ "") +
  geom_hline(yintercept=1, linetype="dashed", color = "blue", size=1.5)+
    labs(x = "", y = "RR")  +
  guides(fill = guide_legend(override.aes = list(size = 3.5))) +
  theme_bw(base_size = 14) +
  theme(strip.text = element_text(face = "bold", size = 13))
f3

```






# Meta-analysis using Maximum likehood Adjusted-----------------------------------------
```{r, echo=FALSE, message=FALSE}
data$TE=log(data$RR)
#dat$seTE=(log(dat$RR_U-log(dat$RR_L))/(3.92))

data$lower <- log(data$RR_L)
data$upper <- log(data$RR_U)
data$seTE <- (data$upper - data$lower)/3.92



l1=is.na(data$RR_L)
dat=data[!l1,]


#dat$TE=log(dat$CRR)
#dat$seTE=(log(dat$CRR_U-log(dat$CRR_L))/(3.92))



m <- metagen(TE,
             seTE,
             data=dat,
             studlab=paste(Study, Country,Year,sep = ", "),
             #studlab=paste(Study),
             comb.fixed = T,
             comb.random = T,
             hakn = F,
             prediction=F,
             sm="RR")



```



#Influence analysis  using leave-one-out -----------------------------------------
```{r}
pdf("figs/InfluencialPlots.pdf", width = 10, height = 20)
forest(metainf(m, pooled = "random"))
dev.off()

inf1=metainf(m, pooled = "random")
summary(inf1)
```

```{r fig.height=4, fig.width=5, echo=FALSE,warning=FALSE}
forest(inf1,
       leftlabs = c("Author, Year of publication, Country", "CMV+", "Number tested"),
       digits = 2,
       col.by="black",
       #sortvar = "TE",
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
      # xlim = c(0, 1), at = c(0,0.5,1),pscale = 100,
       backtransf = T)

```




# trim and fill : Prevalence -----------------------------------------



```{r}
tf2 <- trimfill(m)
summary(tf2)
pdf("figs/Trimmed.pdf", width = 10, height = 10)
funnel(tf2)
dev.off()
funnel(tf2, pch = ifelse(tf2$trimfill, 1, 16),
       level = 0.9, comb.random = T,backtransf = T)
#
```




# Funnel plots : Prevalence -----------------------------------------
```{r}

pdf("figs/FunnelPlotFinal.pdf", width = 10, height = 10)
funnel(m,xlab = "log transformed proportions",studlab = F)
dev.off()

metabias(m, method="linreg")

```

```{r}

metabias(m, method="rank")

```





# Overall effect of steroids on mortality
```{r,fig.height=5, fig.width=10, echo=FALSE,warning=FALSE, message=F, eval=T}


#t<- update(m, byvar=dat$X,print.byvar=F,comb.random = TRUE)
#pdf("figs/Overall.pdf", width = 8, height = 8)
forest(m,sortvar = TE,test.overall = T, overall= T, overall.hetstat = T,bysort = F,subgroup = TRUE,
       
      #colgap.forest.left = "0.2 cm",
       print.byvar = F,
       squaresize=0.7, 
       col.square="blue",
       col.diamond="maroon", 
       col.diamond.lines="maroon",
       #xlim = c(0, 100), at = c(0,5,100),
       print.tau2=F,  
       xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
      col.by="black",
      leftcols=("studlab"),
      leftlabs = c("Study, Country, Year"),
       # xlim = c(0,4),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=F,test.overall.random = F,   print.pval.Q = T,
       digits.pval.Q=4)
#dev.off()

```


# Meta-regression
```{r time1}


#output.metareg=metareg(m,Mean_age)
output.metareg=metareg(m, ~mean_day+Median_age)

#m.gen.reg <- metareg(m.gen, ~year)


meta::bubble(output.metareg,
       xlab = "time (d)",
       col.line = "blue",
       studlab = TRUE)
output.metareg

```



# Study design
```{r,fig.height=20, fig.width=10, echo=FALSE,warning=FALSE}
#sto<- update(m, byvar=dat$Study_Pub,print.byvar=F,comb.random = TRUE)
msd<- update(m, byvar=Study_Design,print.byvar=F)
#m5$bylevs<- c("5 days",">5 days")

pdf("figs/Design.pdf", width = 10, height = 10)
forest(msd,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,
       digits.pval.Q=4)
dev.off()

```



# Steroid type
```{r,fig.height=20, fig.width=10, echo=FALSE,warning=FALSE}
#sto<- update(m, byvar=dat$Study_Pub,print.byvar=F,comb.random = TRUE)
mst<- update(m, byvar=steroid,print.byvar=F)
#m5$bylevs<- c("5 days",">5 days")

pdf("figs/Steroid_Type.pdf", width = 10, height = 12)
forest(mst,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,
       digits.pval.Q=4)
dev.off()

```


# WHO 
```{r,fig.height=20, fig.width=10, echo=FALSE,warning=FALSE}
#sto<- update(m, byvar=dat$Study_Pub,print.byvar=F,comb.random = TRUE)
who<- update(m, byvar=WHO,print.byvar=F)
#m5$bylevs<- c("5 days",">5 days")

pdf("figs/WHO.pdf", width = 10, height = 12)
forest(who,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,
       digits.pval.Q=4)
dev.off()

```






# Only with time -----------------------------------------------------------------------
```{r, echo=FALSE}
data$TE=log(data$RR)
#dat$seTE=(log(dat$RR_U-log(dat$RR_L))/(3.92))

data$lower <- log(data$RR_L)
data$upper <- log(data$RR_U)
data$seTE <- (data$upper - data$lower)/3.92



l1=is.na(data$mean_day)
dat=data[!l1,]


#dat$TE=log(dat$CRR)
#dat$seTE=(log(dat$CRR_U-log(dat$CRR_L))/(3.92))



mt <- metagen(TE,
             seTE,
             data=dat,
            studlab=paste(Study, Country,Year,sep = ", "),
             comb.fixed = T,
             comb.random = T,
             hakn = F,
             prediction=F,
             sm="RR")



```



# Meta-regression
```{r}


#output.metareg=metareg(m,Mean_age)
#output.metareg=metareg(m ~ relevel(factor(Time_6), ref="Time_67 and more days"))

#m.gen.reg <- metareg(m.gen, ~year)


#print(output.metareg, digits=3)
#round(exp(coef(summary(output.metareg))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)



who=metareg(m, duration_symptoms)
who
round(exp(coef(summary(who))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)

```



# 5 days and less
```{r,fig.height=10, fig.width=10, echo=FALSE,warning=FALSE}
#sto<- update(m, byvar=dat$Study_Pub,print.byvar=F,comb.random = TRUE)
m5<- update(mt, byvar=Time_5,print.byvar=F)
m5$bylevs<- c("5 days",">5 days")

pdf("figs/5days.pdf", width = 7, height = 7)
forest(m5,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,
       digits.pval.Q=4)
dev.off()

```



# 6 days and less
```{r,fig.height=10, fig.width= 10, echo=FALSE,warning=FALSE}

m6<- update(mt, byvar=Time_6,print.byvar=F)
m6$bylevs<- c(" 6 days","> 6 days")
pdf("figs/6days.pdf", width = 7, height = 7)
forest(m6,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study"),
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T, print.tau2=F,  print.pval.Q = T,
       digits.pval.Q=4)
dev.off()

```


# 7 days and less
```{r,fig.height=10, fig.width=10, echo=FALSE,warning=FALSE}

m7<- update(mt, byvar=Time_7,print.byvar=F)
m7$bylevs<- c("7 days","> 7 days")
pdf("figs/7days.pdf", width = 7, height = 7)
forest(m7,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.2 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study"),
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T, print.tau2=F,  print.pval.Q = T,
       digits.pval.Q=4)
dev.off()

```



```{r,fig.height=15, fig.width=15, echo=FALSE,warning=FALSE, eval=F}

#pdf("figs/Fig1pdf", width = 15, height = 30)
forest(mnds,
       leftlabs = c("Author", "Died (n)","Non-Steroid (N)"),
       digits = 2,
       col.by="black",
       sortvar = TE,
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       xlab="Death in non-steroid group (%)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(0, 100), at = c(0,10, 20,30,40,50,60,70,80,90, 100),pscale = 100,
       backtransf = T)
#dev.off()

```
```{r, eval=F, echo=F}
#dat <- read_xls(file= "steroids3_NVedit.xlsx", header=T,stringsAsFactors=T)
dat <- read.csv("steroids5.csv")


res <- rma(ai=dat$Steroids_died, bi=dat$Steroid_Alive, ci=dat$No_Steroid_Died, di=dat$No_Steroid_Alive, data=dat, measure="RR",
           slab=paste(dat$Study, sep = ","), method="DL")


#tiff(file="Odds Ratio_CVD.tif",bg="white", width = 1500, height = 1024, units = "px",pointsize = 14)

pdf("OR_steroids.pdf", width = 20, height = 15)


forest(res, transf=exp, xlim=c(-6000,1000), addcred=F, showweights=F, refline=1, xlab="Odds Ratio (95%CI)", 
       mlab="Random Effect Model",cex=1.5,at=c(0,100),ylim=c(-1, 40),
       ilab=cbind(ai=dat$Steroids_died, bi=dat$Steroid_Alive, ci=dat$No_Steroid_Died, di=dat$No_Steroid_Alive),
       ilab.xpos=c(-3500,-2600, -1600,-700))

par(cex=.75, font=2)
text(c(-3500,-2600,-1600,-700), 35, c("Mortality+", "Mortality-", "Mortality+", "Mortality-"),cex=1.8)
text(c(-3200,-1000),     35, c("Steroids=Yes", "Steroids=No"),cex=1.9)
text(-6000,                33, "Author(s), Ending Year of Study",   pos=4,cex=1.8)
text(1000,                  30, "Odds Ratio [95% CI]", pos=2,cex=1.8)


dev.off()

```


```{r}

summary(data$Median_age)
table(data$Study_Design)
sum(data$Sample_size, na.rm = T)
table(data$Country)
table(data$steroid)
```

# RCT.............................................................................................................
```{r}
dat=read.csv("steroids7.csv")

dat<- dat[ which(dat$Study.design=="RCT"), ]


# Meta-analysis using Maximum likehood Adjusted-----------------------------------------

dat$TE=log(dat$RR)
#dat$seTE=(log(dat$RR_U-log(dat$RR_L))/(3.92))

dat$lower <- log(dat$RR_L)
dat$upper <- log(dat$RR_U)
dat$seTE <- (dat$upper - dat$lower)/3.92



l1=is.na(dat$RR_L)
dat=dat[!l1,]


#dat$TE=log(dat$CRR)
#dat$seTE=(log(dat$CRR_U-log(dat$CRR_L))/(3.92))



m <- metagen(TE,
             seTE,
             data=dat,
             studlab=paste(Study, Country,Year,sep = ", "),
             #studlab=paste(Study),
             comb.fixed = T,
             comb.random = T,
             hakn = F,
             prediction=F,
             sm="RR")

```

# Remdesivir
```{r,fig.height=4, fig.width=5, echo=FALSE,warning=FALSE}
#sto<- update(m, byvar=dat$Study_Pub,print.byvar=F,comb.random = TRUE)
msd<- update(m, byvar=dat$Remdesivir,print.byvar=T)
#m5$bylevs<- c("5 days",">5 days")

#pdf("figs/Design.pdf", width = 10, height = 10)
forest(msd,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = T,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,
       digits.pval.Q=4)
#dev.off()

```


# Predefined mortality
```{r,fig.height=4, fig.width=5, echo=FALSE,warning=FALSE}
#sto<- update(m, byvar=dat$Study_Pub,print.byvar=F,comb.random = TRUE)
msd<- update(m, byvar=dat$Predefined_mortality,print.byvar=T)
#m5$bylevs<- c("5 days",">5 days")

#pdf("figs/Design.pdf", width = 10, height = 10)
forest(msd,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = T,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,
       digits.pval.Q=4)
#dev.off()

```


# Remdesivir
```{r,fig.height=4, fig.width=5, echo=FALSE,warning=FALSE}
#sto<- update(m, byvar=dat$Study_Pub,print.byvar=F,comb.random = TRUE)
msd<- update(m, byvar=dat$Remdesivir,print.byvar=T)
#m5$bylevs<- c("5 days",">5 days")

#pdf("figs/Design.pdf", width = 10, height = 10)
forest(msd,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = T,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,
       digits.pval.Q=4)
#dev.off()

```



# Prop Oxygen

```{r Oxygen}

output.metareg5=metareg(m,Oxygen)
meta::bubble(output.metareg5,
       xlab = "Prop Oxygen (%)",
       col.line = "blue",
       studlab = TRUE)
m$seTE


```



# Prop Mechanical Ventilation

```{r MV}

output.metareg5=metareg(m,MV)
meta::bubble(output.metareg5,
       xlab = "Prop Mechanical Ventilation (%)",
       col.line = "blue",
       studlab = TRUE)
m$seTE


```




#Durations of symptoms

```{r Symptoms}

output.metareg5=metareg(m, duration_symptoms)
meta::bubble(output.metareg5,
       xlab = "Duration of symptoms (d)",
       col.line = "blue",
       studlab = TRUE)
m$seTE

```



# Control steroids

```{r}

output.metareg5=metareg(m, Control_Steroid)
meta::bubble(output.metareg5,
       xlab = "Steroids in the control group",
       col.line = "blue",
       studlab = TRUE)

output.metareg5

```

# Meta-regression
```{r multivariable}


#output.metareg=metareg(m,Mean_age)
output.metareg=metareg(m, ~Time_6)

#m.gen.reg <- metareg(m.gen, ~year)


meta::bubble(output.metareg,
       xlab = "time (d)",
       col.line = "blue",
       studlab = TRUE)


print(output.metareg, digits=3)
round(exp(coef(summary(output.metareg))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)

```


```{r}
m.qual.rep <- rma(yi = TE, 
                  sei = seTE, 
                  data = m, 
                  method = "ML", 
                  mods = ~ Time_6, 
                  test = "knha")

m.qual.rep
```


```{r}

res <- rma(yi, vi,mods = ~ relevel(factor(Specimen_Screening2), ref="Urine or Saliva"),dat=dat3)
print(res, digits=3)
round(exp(coef(summary(res))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)

```



