---
title: "Optimal Duration of Systemic Steroids Use in COVID-19 Treatment: A systematic Review and Meta-analysis"

author: ''
date: "02/01/2023"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE,warning=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE,echo=FALSE}
library(meta)
library(metafor)
```


```{r, echo=FALSE}

rm(list=ls())
data=read.csv("steroids7.csv")
```

```{r}
sum(data$Sample_size)
median (data$Mean_or_Median_Age, na.rm = T)
median (data$X..Male, na.rm = T)
table(data$Study_Design7)
table(data$steroid)
table(data$Country)
table(data$WHO)

```


# Meta-analysis using Maximum likehood Adjusted-----------------------------------------
```{r, echo=FALSE, message=FALSE}
data$TE=log(data$RR)

data$lower <- log(data$RR_L)
data$upper <- log(data$RR_U)
data$seTE <- (data$upper - data$lower)/3.92

l1=is.na(data$RR_L)
dat=data[!l1,]


m <- metagen(TE,
             seTE,
             data=dat,
             studlab=paste(Study, Country,Year,sep = ", "),
             #studlab=paste(Study),
             comb.fixed = F,
             comb.random = T,
             hakn = F,
             prediction=F,
             sm="RR")

```



#Influence analysis  using leave-one-out -----------------------------------------
```{r}
pdf("figs/InfluencialPlots.pdf", width = 10, height = 10)
forest(metainf(m, pooled = "random"))
dev.off()

inf1=metainf(m, pooled = "random")
summary(inf1)
```

# trim and fill :-----------------------------------------

```{r}
tf2 <- trimfill(m)
summary(tf2)
pdf("figs/Trimmed.pdf", width = 10, height = 10)
funnel(tf2)
dev.off()
funnel(tf2, pch = ifelse(tf2$trimfill, 1, 16),
       level = 0.9, comb.random = T,backtransf = T)
```

# Funnel plots :-----------------------------------------
```{r}

pdf("figs/FunnelPlotFinal.pdf", width = 10, height = 10)
funnel(m,xlab = "log transformed proportions",studlab = F)
dev.off()

metabias(m, method="linreg")

```

```{r}

metabias(m, method="rank")

```


# Overall effect of steroids on mortality
```{r,fig.height=5, fig.width=10, echo=FALSE,warning=FALSE, message=F, eval=T}

tiff("figs/Overall.TIFF", units="in", width=10, height=10, res=300)
forest(m,sortvar = TE,test.overall = T, overall= T, overall.hetstat = T,bysort = F,subgroup = TRUE,
       
      #colgap.forest.left = "0.2 cm",
       print.byvar = F,
       squaresize=0.7, 
       col.square="blue",
       col.diamond="maroon", 
       col.diamond.lines="maroon",
       #xlim = c(0, 100), at = c(0,5,100),
       print.tau2=F,  
       xlab="Association between Systemic Steroid and COVID-19 Mortality",
      col.by="black",
      leftcols=("studlab"),
      leftlabs = c("Study, Country, Year"),
       # xlim = c(0,4),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=F,test.overall.random = F,   print.pval.Q = T,test.subgroup.random =F,
       digits.pval.Q=4)
dev.off()

```


# Meta-regression
```{r time1}


#output.metareg=metareg(m,Mean_age)
output.metareg=metareg(m, ~mean_day+Median_age)

#m.gen.reg <- metareg(m.gen, ~year)


meta::bubble(output.metareg,
       xlab = "time (d)",
       col.line = "blue",
       studlab = TRUE)
output.metareg

```



# Study design
```{r,fig.height=20, fig.width=10, echo=FALSE,warning=FALSE}
msd<- update(m, byvar=Study_Design7,print.byvar=F)

tiff("figs/Design.TIFF", units="in", width=10, height=10, res=300)
forest(msd,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,test.subgroup.random =F,
       digits.pval.Q=4)
dev.off()

```


# Steroid type
```{r,fig.height=20, fig.width=10, echo=FALSE,warning=FALSE}
mst<- update(m, byvar=steroid,print.byvar=F)

tiff("figs/Steroid_Type.TIFF", units="in", width=10, height=10, res=300)
forest(mst,sortvar = TE,test.overall = F, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=F,test.overall.random = T,print.pval.Q = F,test.subgroup.random =F,
       digits.pval.Q=4)
dev.off()

```


# WHO 
```{r,fig.height=20, fig.width=10, echo=FALSE,warning=FALSE}
who<- update(m, byvar=WHO,print.byvar=F)

tiff("figs/WHO.TIFF", units="in", width=10, height=10, res=300)
forest(who,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       print.tau2=F,
        xlab="Association between Systemic Steroid Use and COVID-19 Mortality",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study, Country, Year"),
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T,print.pval.Q = T,test.subgroup.random =F,
       digits.pval.Q=4)
dev.off()

```






# Only with time -----------------------------------------------------------------------
```{r, echo=FALSE}
data$TE=log(data$RR)

data$lower <- log(data$RR_L)
data$upper <- log(data$RR_U)
data$seTE <- (data$upper - data$lower)/3.92

l1=is.na(data$mean_day)
dat=data[!l1,]


mt <- metagen(TE,
             seTE,
             data=dat,
            studlab=paste(Study, Country,Year,sep = ", "),
             comb.fixed = F,
             comb.random = T,
             hakn = F,
             prediction=F,
             sm="RR")



```



# Meta-regression
```{r}


output.metareg=metareg(m,Mean_age)
output.metareg=metareg(m ~ relevel(factor(Time_6), ref="Time_67 and more days"))

m.gen.reg <- metareg(m.gen, ~year)


print(output.metareg, digits=3)
round(exp(coef(summary(output.metareg))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)



who=metareg(m, duration_symptoms)
who
round(exp(coef(summary(who))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)

```



# 5 days and less
```{r,fig.height=10, fig.width=10, echo=FALSE,warning=FALSE}
#sto<- update(m, byvar=dat$Study_Pub,print.byvar=F,comb.random = TRUE)
m5<- update(mt, byvar=Time_5,print.byvar=F)
m5$bylevs<- c("5 days",">5 days")

#pdf("figs/5days.pdf", width = 7, height = 7)
tiff("figs/5days.TIFF", units="in", width=10, height=10, res=300)
forest(m5,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study"),
        xlab="Association between Systemic Steroid and COVID-19 Mortality",
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T, print.tau2=F,  print.pval.Q = T,test.subgroup.random =F,
       digits.pval.Q=4)
dev.off()

```



# 6 days and less
```{r,fig.height=10, fig.width= 10, echo=FALSE,warning=FALSE}

m6<- update(mt, byvar=Time_6,print.byvar=F)
m6$bylevs<- c(" 6 days","> 6 days")
#pdf("figs/6days.pdf", width = 7, height = 7)
tiff("figs/6days.TIFF", units="in", width=10, height=10, res=300)

forest(m6,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study"),
        xlab="Association between Systemic Steroid and COVID-19 Mortality",
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T, print.tau2=F,  print.pval.Q = T,test.subgroup.random =F,
       digits.pval.Q=4)
dev.off()
m6
d6=metareg(mt,Time_6)
d6
round(exp(coef(summary(d6))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)

```


# 7 days and less
```{r,fig.height=10, fig.width=10, echo=FALSE,warning=FALSE}

m7<- update(mt, byvar=Time_7,print.byvar=F)
m7$bylevs<- c("7 days","> 7 days")
tiff("figs/7days.TIFF", units="in", width=10, height=10, res=300)

forest(m7,sortvar = TE,test.overall = T, overall= F, overall.hetstat = F,
       colgap.forest.left = "0.38 cm",print.byvar = F,
       squaresize=0.5, col.square="blue",
       col.diamond="maroon", col.diamond.lines="maroon",
       col.by="black",leftcols=("studlab"),leftlabs = c("Study"),
        xlab="Association between Systemic Steroid and COVID-19 Mortality",
       rightlabs = c("RR","95% CI","Weight"), just = "center", just.addcols = "center",
       print.zval = FALSE,text.fixed="Overall (Fixed-Effect Model)",text.random="Overall (Random-Effect Model)",comb.random = T,comb.fixed = F,
       test.overall.fixed=T,test.overall.random = T, print.tau2=F,  print.pval.Q = T,test.subgroup.random =F,
       digits.pval.Q=4)
dev.off()

```

```{r}

summary(data$Median_age)
table(data$Study_Design)
sum(data$Sample_size, na.rm = T)
table(data$Country)
table(data$steroid)
```

# RCT.............................................................................................................
```{r}
dat=read.csv("steroids7.csv")

dat<- dat[ which(dat$Study.design=="RCT"), ]


# Meta-analysis using Maximum likehood Adjusted-----------------------------------------

dat$TE=log(dat$RR)
#dat$seTE=(log(dat$RR_U-log(dat$RR_L))/(3.92))

dat$lower <- log(dat$RR_L)
dat$upper <- log(dat$RR_U)
dat$seTE <- (dat$upper - dat$lower)/3.92



l1=is.na(dat$RR_L)
dat=dat[!l1,]


#dat$TE=log(dat$CRR)
#dat$seTE=(log(dat$CRR_U-log(dat$CRR_L))/(3.92))



m <- metagen(TE,
             seTE,
             data=dat,
             studlab=paste(Study, Country,Year,sep = ", "),
             #studlab=paste(Study),
             comb.fixed = F,
             comb.random = T,
             hakn = F,
             prediction=F,
             sm="RR")
m
```


# Prop Oxygen

```{r Oxygen}

output.metareg5=metareg(m,Oxygen)
meta::bubble(output.metareg5,
       xlab = "Prop Oxygen (%)",
       col.line = "blue",
       studlab = TRUE)
m$seTE

round(exp(coef(summary(output.metareg5))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)


```



# Prop Mechanical Ventilation

```{r MV}

output.metareg5=metareg(m,MV)
meta::bubble(output.metareg5,
       xlab = "Prop Mechanical Ventilation (%)",
       col.line = "blue",
       studlab = TRUE)

output.metareg5

round(exp(coef(summary(output.metareg5))[-1,c("estimate", "ci.lb", "ci.ub")]), 4)
```


#Durations of symptoms

```{r Symptoms}

output.metareg5=metareg(m, duration_symptoms)
meta::bubble(output.metareg5,
       xlab = "Duration of symptoms (d)",
       col.line = "blue",
       studlab = TRUE)

output.metareg5

round(exp(coef(summary(output.metareg5))[-1,c("estimate", "ci.lb", "ci.ub")]), 4)


```


# Control steroids

```{r}

output.metareg5=metareg(m, Control_Steroid)
meta::bubble(output.metareg5,
       xlab = "Steroids in the control group",
       col.line = "blue",
       studlab = TRUE)

output.metareg5

round(exp(coef(summary(output.metareg5))[-1,c("estimate", "ci.lb", "ci.ub")]), 4)

```

# Meta-regression
```{r multivariable}


#output.metareg=metareg(m,Mean_age)
output.metareg=metareg(m, ~Time_6)

#m.gen.reg <- metareg(m.gen, ~year)


meta::bubble(output.metareg,
       xlab = "time (d)",
       col.line = "blue",
       studlab = TRUE)


print(output.metareg, digits=3)
round(exp(coef(summary(output.metareg))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)

```


# Age

```{r}

output.metareg5=metareg(m, Median_age)
meta::bubble(output.metareg5,
       xlab = "Age",
       col.line = "blue",
       studlab = TRUE)

output.metareg5

round(exp(coef(summary(output.metareg5))[-1,c("estimate", "ci.lb", "ci.ub")]), 4)


```





